---
title: "Correlation"
output: html_notebook
---

## Load packages
```{r}
library(tidyverse)
library(readxl)
library(ggpubr)
library(broom)
```

## Load data
```{r}
# Rt data
rt <- read_csv('estR/estR.csv') %>% 
  select(`Mean(R)`, date_start, Region) %>% 
  rename(Rt=`Mean(R)`, Date=date_start)
t_rt <- rt %>% filter(Region == 'Toronto')
p_rt <- rt %>% filter(Region == 'Peel')
y_rt <- rt %>% filter(Region == 'York')
h_rt <- rt %>% filter(Region == 'Halton')
d_rt <- rt %>% filter(Region == 'Durham')

# Mobility data
regions = c('Regional Municipality of Durham', 'Regional Municipality of Halton', 'Regional Municipality of Peel', 'Toronto Division', 'Regional Municipality of York')
mob <- read_csv('data/2020_CA_Region_Mobility_Report.csv')
mob <- mob %>% 
  mutate(`Global Mobility Score` = (retail_and_recreation_percent_change_from_baseline + grocery_and_pharmacy_percent_change_from_baseline + parks_percent_change_from_baseline + transit_stations_percent_change_from_baseline + workplaces_percent_change_from_baseline + residential_percent_change_from_baseline)/6)
t_mob <- mob %>% 
  filter(sub_region_2 == 'Toronto Division')
p_mob <- mob %>% 
  filter(sub_region_2=='Regional Municipality of Peel')
y_mob <- mob %>% 
  filter(sub_region_2=='Regional Municipality of York')
d_mob <- mob %>% 
  filter(sub_region_2=='Regional Municipality of Durham')
h_mob <- mob %>% 
  filter(sub_region_2=='Regional Municipality of Halton')
```

## Function to convert data into useful form
Input: mobility data, rt data, lockdown start date, lockdown end date, columns to keep  
Output: Df of specific region to run regressions on
```{r}
get_data <- function(mob_data, rt_data, lock_start, lock_end, cols) {
  df <- mob_data %>% 
    left_join(rt_data, by=c("date" = "Date")) %>% 
    select(all_of(cols)) %>% 
    mutate(date = as.Date(date)) %>% 
    rename(retail=retail_and_recreation_percent_change_from_baseline,grocery=grocery_and_pharmacy_percent_change_from_baseline, parks=parks_percent_change_from_baseline, transit=transit_stations_percent_change_from_baseline, workplaces=workplaces_percent_change_from_baseline, residential=residential_percent_change_from_baseline, global=`Global Mobility Score`) %>% 
    mutate(retail_lag7 = lag(retail, 7)) %>% 
    mutate(grocery_lag7 = lag(grocery, 7) ) %>% 
    mutate(parks_lag7 = lag(parks, 7)) %>%
    mutate(transit_lag7 = lag(transit, 7)) %>%
    mutate(workplaces_lag7 = lag(workplaces, 7)) %>%
    mutate(residential_lag7 = lag(residential, 7)) %>%
    mutate(global_lag7 = lag(global, 7)) %>%
    mutate(retail_lag14 = lag(retail, 14)) %>% 
    mutate(grocery_lag14 = lag(grocery, 14) ) %>% 
    mutate(parks_lag14 = lag(parks, 14)) %>%
    mutate(transit_lag14 = lag(transit, 14)) %>%
    mutate(workplaces_lag14 = lag(workplaces, 14)) %>%
    mutate(residential_lag14 = lag(residential, 14)) %>%
    mutate(global_lag14 = lag(global, 14)) %>%
    filter(date>=lock_start & date<=lock_end)
  return(df)
}
```

## Function to get correlation data
Input: Df of specific region, desired mobility lag time (0, 7, or 14 days)  
Output: Df of specific region with list of correlations for each mobility
```{r}
get_correlations <- function(data, lag) {
  if(lag==0) {
    retail <- cor.test(data$retail, data$Rt, method = 'pearson')
    grocery <- cor.test(data$grocery, data$Rt, method = 'pearson')
    parks <- cor.test(data$parks, data$Rt, method = 'pearson')
    transit <- cor.test(data$transit, data$Rt, method = 'pearson')
    workplaces <- cor.test(data$workplaces, data$Rt, method = 'pearson')
    residential <- cor.test(data$residential, data$Rt, method = 'pearson')
    global <- cor.test(data$global, data$Rt, method = 'pearson')
  } else if(lag==7) {
    retail <- cor.test(data$retail_lag7, data$Rt, method = 'pearson')
    grocery <- cor.test(data$grocery_lag7, data$Rt, method = 'pearson')
    parks <- cor.test(data$parks_lag7, data$Rt, method = 'pearson')
    transit <- cor.test(data$transit_lag7, data$Rt, method = 'pearson')
    workplaces <- cor.test(data$workplaces_lag7, data$Rt, method = 'pearson')
    residential <- cor.test(data$residential_lag7, data$Rt, method = 'pearson')
    global <- cor.test(data$global_lag7, data$Rt, method = 'pearson')
  } else if(lag==14) {
    retail <- cor.test(data$retail_lag14, data$Rt, method = 'pearson')
    grocery <- cor.test(data$grocery_lag14, data$Rt, method = 'pearson')
    parks <- cor.test(data$parks_lag14, data$Rt, method = 'pearson')
    transit <- cor.test(data$transit_lag14, data$Rt, method = 'pearson')
    workplaces <- cor.test(data$workplaces_lag14, data$Rt, method = 'pearson')
    residential <- cor.test(data$residential_lag14, data$Rt, method = 'pearson')
    global <- cor.test(data$global_lag14, data$Rt, method = 'pearson')
  } else {
    stop('Must enter 0, 7, or 14')
  }

  return(list(Retail=retail, Grocery=grocery, Parks=parks, Transit=transit, Workplaces=workplaces, Residential=residential, Global=global))
}
```

## Function to get correlation coefficient
Input: List of region df's each containing coefficient info  
Output: Coefficient data df of specified lag time
```{r}
corr_coef <- function(data) {
  place <- numeric(length=35)
  mobility <- numeric(length=35)
  coefficient <- numeric(length=35)
  p_val <- numeric(length=35)
  i=1
  j=1
  k=1
  for(loc in data){
    for(corr in loc) {
      place[i] <- names(data)[j]
      mobility[i] <- names(loc)[k]
      coefficient[i] <- corr$estimate
      p_val[i] <- corr$p.value
      i=i+1
      k=k+1
    }
    j=j+1
    k=1
  }
  df <- data.frame(place, mobility, coefficient, p_val)
  return(df)
}
```

## Function to get regression data
Input: Df of specific region, desired mobility lag time (0, 7, or 14 days)  
Output: Df of specific region with list of regressions for each mobility
```{r}
get_regressions <- function(data, lag) {
  if(lag==0) {
    retail <- glm(Rt ~ retail, data = data)
    grocery <- glm(Rt ~ grocery, data = data)
    parks <- glm(Rt ~ parks, data = data)
    transit <- glm(Rt ~ transit, data = data)
    workplaces <- glm(Rt ~ workplaces, data = data)
    residential <- glm(Rt ~ residential, data = data)
    global <- glm(Rt ~ global, data = data)
  } else if(lag==7) {
    retail <- glm(Rt ~ retail_lag7, data = data)
    grocery <- glm(Rt ~ grocery_lag7, data = data)
    parks <- glm(Rt ~ parks_lag7, data = data)
    transit <- glm(Rt ~ transit_lag7, data = data)
    workplaces <- glm(Rt ~ workplaces_lag7, data = data)
    residential <- glm(Rt ~ residential_lag7, data = data)
    global <- glm(Rt ~ global_lag7, data = data)
  } else if(lag==14) {
    retail <- glm(Rt ~ retail_lag14, data = data)
    grocery <- glm(Rt ~ grocery_lag14, data = data)
    parks <- glm(Rt ~ parks_lag14, data = data)
    transit <- glm(Rt ~ transit_lag14, data = data)
    workplaces <- glm(Rt ~ workplaces_lag14, data = data)
    residential <- glm(Rt ~ residential_lag14, data = data)
    global <- glm(Rt ~ global_lag14, data = data)
  } else {
    stop('Must enter 0, 7, or 14')
  }
  return(list(Retail=retail, Grocery=grocery, Parks=parks, Transit=transit, Workplaces=workplaces, Residential=residential, Global=global))
}
```

## Function to get regression coefficient
Input: List of region df's each containing regression info  
Output: Coefficient data df of specified lag time
```{r}
reg_coef <- function(data) {
  place <- numeric(length=35)
  mobility<- numeric(length=35)
  coefficient <- numeric(length=35)
  r2 <- numeric(length=35)
  p_val <- numeric(length=35)
  i=1
  j=1
  k=1
  for(loc in data){
    for(reg in loc) {
      place[i] <- names(data)[j]
      mobility[i] <- names(loc)[k]
      coefficient[i] <- tidy(reg)[[2,2]]
      # r2[i] <- glance(reg)[[1,2]]
      p_val[i] <- tidy(reg)[[2,5]]
      i=i+1
      k=k+1
    }
    j=j+1
    k=1
  }
  df <- data.frame(place, mobility, coefficient, p_val)
  return(df)
}
```

## Pipeline
Columns to keep
```{r}
COLS <- c('sub_region_2', 'date', 'retail_and_recreation_percent_change_from_baseline','grocery_and_pharmacy_percent_change_from_baseline', 'parks_percent_change_from_baseline', 'transit_stations_percent_change_from_baseline', 'workplaces_percent_change_from_baseline', 'residential_percent_change_from_baseline', 'Global Mobility Score', 'Rt')
```

Get data
```{r}
# Wave 1
d <- get_data(d_mob, d_rt, '2020-03-17', '2020-06-19', COLS)
p <- get_data(p_mob, p_rt, '2020-03-17', '2020-06-24', COLS)
t <- get_data(t_mob, t_rt, '2020-03-17', '2020-06-24', COLS)
h <- get_data(h_mob, h_rt, '2020-03-17', '2020-06-19', COLS)
y <- get_data(y_mob, y_rt, '2020-03-17', '2020-06-19', COLS)

# Wave 2
d2 <- get_data(d_mob, d_rt, '2020-12-26', '2021-02-16', COLS)
p2 <- get_data(p_mob, p_rt, '2020-12-26', '2021-03-19', COLS)
t2 <- get_data(t_mob, t_rt, '2020-12-26', '2021-03-19', COLS)
h2 <- get_data(h_mob, h_rt, '2020-12-26', '2021-02-16', COLS)
y2 <- get_data(y_mob, y_rt, '2020-12-26', '2021-02-22', COLS)
```

0 day lag
```{r}
# Wave 1
zero_corr <- list(
  durham=get_correlations(d, 0),
  peel=get_correlations(p, 0),
  toronto=get_correlations(t, 0),
  halton=get_correlations(h, 0),
  york=get_correlations(y, 0)
)
zero_reg <- list(
  durham=get_regressions(d, 0),
  peel=get_regressions(p, 0),
  toronto=get_regressions(t, 0),
  halton=get_regressions(h, 0),
  york=get_regressions(y, 0)
)

# Wave 2
zero_corr2 <- list(
  durham=get_correlations(d2, 0),
  peel=get_correlations(p2, 0),
  toronto=get_correlations(t2, 0),
  halton=get_correlations(h2, 0),
  york=get_correlations(y2, 0)
)
zero_reg2 <- list(
  durham=get_regressions(d2, 0),
  peel=get_regressions(p2, 0),
  toronto=get_regressions(t2, 0),
  halton=get_regressions(h2, 0),
  york=get_regressions(y2, 0)
)
```

7 day lag
```{r}
# Wave 1
seven_corr <- list(
  durham=get_correlations(d, 7),
  peel=get_correlations(p, 7),
  toronto=get_correlations(t, 7),
  halton=get_correlations(h, 7),
  york=get_correlations(y, 7)
)
seven_reg <- list(
  durham=get_regressions(d, 7),
  peel=get_regressions(p, 7),
  toronto=get_regressions(t, 7),
  halton=get_regressions(h, 7),
  york=get_regressions(y, 7)
)

# Wave 2
seven_corr2 <- list(
  durham=get_correlations(d2, 7),
  peel=get_correlations(p2, 7),
  toronto=get_correlations(t2, 7),
  halton=get_correlations(h2, 7),
  york=get_correlations(y2, 7)
)
seven_reg2 <- list(
  durham=get_regressions(d2, 7),
  peel=get_regressions(p2, 7),
  toronto=get_regressions(t2, 7),
  halton=get_regressions(h2, 7),
  york=get_regressions(y2, 7)
)
```

14 day lag
```{r}
fourteen_corr <- list(
  durham = get_correlations(d, 14),
  peel = get_correlations(p, 14),
  toronto = get_correlations(t, 14),
  halton = get_correlations(h, 14),
  york = get_correlations(y, 14)
)
fourteen_reg <- list(
  durham = get_regressions(d, 14),
  peel = get_regressions(p, 14),
  toronto = get_regressions(t, 14),
  halton = get_regressions(h, 14),
  york = get_regressions(y, 14)
)
fourteen_corr2 <- list(
  durham = get_correlations(d2, 14),
  peel = get_correlations(p2, 14),
  toronto = get_correlations(t2, 14),
  halton = get_correlations(h2, 14),
  york = get_correlations(y2, 14)
)
fourteen_reg2 <- list(
  durham = get_regressions(d2, 14),
  peel = get_regressions(p2, 14),
  toronto = get_regressions(t2, 14),
  halton = get_regressions(h2, 14),
  york = get_regressions(y2, 14)
)
```

Get coefficient data for each time lag
Add period column, later factor it appropriately
Combine into one df (total)
Check if each p-value is statistically significant (<0.01)
```{r}
# Wave 1
zero_corr_coef <- corr_coef(zero_corr)
zero_corr_coef$period <- rep('Lockdown', times=35)

seven_corr_coef <- corr_coef(seven_corr)
seven_corr_coef$period <- rep('7 Days after lockdown', times=35)

fourteen_corr_coef <- corr_coef(fourteen_corr)
fourteen_corr_coef$period <- rep('14 Days after lockdown', times=35)

total_corr <- rbind(zero_corr_coef, seven_corr_coef, fourteen_corr_coef)
total_corr$wave <- rep('Wave 1', times=105)

# Wave 2
zero_corr_coef2 <- corr_coef(zero_corr2)
zero_corr_coef2$period <- rep('Lockdown', times=35)

seven_corr_coef2 <- corr_coef(seven_corr2)
seven_corr_coef2$period <- rep('7 Days after lockdown', times=35)

fourteen_corr_coef2 <- corr_coef(fourteen_corr2)
fourteen_corr_coef2$period <- rep('14 Days after lockdown', times=35)

total_corr2 <- rbind(zero_corr_coef2, seven_corr_coef2, fourteen_corr_coef2)
total_corr2$wave <- rep('Wave 2', times=105)

# Combine
total_correlation <- rbind(total_corr, total_corr2)

total_correlation$period <- factor(total_correlation$period, levels=c('Lockdown', '7 Days after lockdown', '14 Days after lockdown'))
total_correlation <- total_correlation %>%
  mutate(
    `Significant (p<0.01)` = case_when(
      p_val < 0.01 ~ 'Yes',
      p_val >= 0.01 ~ 'No'
    )
  )

write_csv(total_correlation, 'data/corr_coef.csv')
```

Get coefficient data for each time lag
Add period column, later factor it appropriately
Combine into one df (total)
Check if each p-value is statistically significant (<0.01)
```{r}
# Wave 1
zero_reg_coef <- reg_coef(zero_reg)
zero_reg_coef$period <- rep('Lockdown', times=35)

seven_reg_coef <- reg_coef(seven_reg)
seven_reg_coef$period <- rep('7 Days after lockdown', times=35)

fourteen_reg_coef <- reg_coef(fourteen_reg)
fourteen_reg_coef$period <- rep('14 Days after lockdown', times=35)

total_reg <- rbind(zero_reg_coef, seven_reg_coef, fourteen_reg_coef)
total_reg$wave <- rep('Wave 1', times=105)

# Wave 2
zero_reg_coef2 <- reg_coef(zero_reg2)
zero_reg_coef2$period <- rep('Lockdown', times=35)

seven_reg_coef2 <- reg_coef(seven_reg2)
seven_reg_coef2$period <- rep('7 Days after lockdown', times=35)

fourteen_reg_coef2 <- reg_coef(fourteen_reg2)
fourteen_reg_coef2$period <- rep('14 Days after lockdown', times=35)

total_reg2 <- rbind(zero_reg_coef2, seven_reg_coef2, fourteen_reg_coef2)
total_reg2$wave <- rep('Wave 2', times=105)

# Combine
total_regression <- rbind(total_reg, total_reg2)

total_regression$period <- factor(total_regression$period, levels=c('Lockdown', '7 Days after lockdown', '14 Days after lockdown'))
total_regression <- total_regression %>%
  mutate(
    `Significant (p<0.01)` = case_when(
      p_val < 0.01 ~ 'Yes',
      p_val >= 0.01 ~ 'No'
    )
  )

write_csv(total_regression, 'data/reg_coef.csv')
```

## Barplot
```{r}
col <- c('Grocery'='red', 'Parks'='blue', 'Residential'='cyan', 'Retail'='darkgreen', 'Transit'='pink', 'Workplaces'='seagreen1', 'Global'='brown')

p <- ggplot(total_regression, aes(x=place, y=coefficient, fill=fct_relevel(mobility, "Global", after = Inf))) +
  geom_col(position = 'dodge') +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5)) +
  scale_fill_manual(values=col, 
                    name='Mobility type') +
  scale_x_discrete(labels = c('Durham', 'Halton', 'Peel', 'Toronto', 'York')) +
  facet_grid(rows=vars(period), cols=vars(wave)) +
  xlab('Public Health Unit') +
  ylab('Regression coefficient') +
  theme_bw()
ggsave('graphs/reg_barplot.pdf',
       p,
       height = 8.5,
       width = 11,
       units = 'in')
p
```
Regression coefficients of each type of mobility with Rt as the dependent variable. A simple way to grasp regression coefficients is to picture them as linear slopes. There is a strong, positive association between almost each type of mobility score and Rt during the periods 7 days after lockdown and 14 days after lockdown. Most of the time, when mobility score is high, Rt is also high.The exception is residential mobility; there is a strong, negative association between residential mobility score and Rt during the periods 7 days after lockdown and 14 days after lockdown. When residential mobility score is high, Rt is low. This suggests that there is a lag time before changes in global mobility score are reflected in Rt values.


## Lineplot
```{r}
d <- filter(total_correlation, mobility=='Global')
p <- ggplot(d, aes(x=period, y=coefficient, group=place, color=place)) +
  geom_point(d, mapping=aes(shape=`Significant (p<0.01)`), size=3) +
  geom_line(linetype='dashed') +
  scale_color_brewer(palette = 'Set1',
                     name = 'Public Health Unit',
                     breaks=c('durham', 'halton', 'peel', 'toronto', 'york'),
                     labels=c('Durham', 'Halton', 'Peel', 'Toronto', 'York')) +
  facet_grid(cols=vars(wave)) +
  xlab('Period') +
  ylab('Correlation coefficient') +
  theme_bw()
ggsave('graphs/corr_lag_lineplot.pdf',
       p,
       height = 8.5,
       width = 11,
       units = 'in')
p
```
Correlation coefficients between Rt and global mobility score. There is usually a strong, positive correlation between global mobility score and Rt during the periods 7 days after lockdown and 14 days after lockdown. When mobility score is high, Rt is also high. This also suggests that there is a lag time before changes in global mobility score are reflected in Rt values.

Correlation coefficient formulas are used to find how strong a relationship is between data. The formulas return a value between -1 and 1, where:

1 indicates a strong positive relationship.
-1 indicates a strong negative relationship.
A result of zero indicates no relationship at all.

```{r}
d <- filter(total_correlation, mobility=='Global')
p <- ggplot(d, aes(x=place, y=coefficient, group=period, color=period)) +
  geom_point(d, mapping=aes(shape=`Significant (p<0.01)`), size=3) +
  geom_line() +
  scale_color_discrete(name = "Period") +
  scale_x_discrete(labels = c('Durham', 'Halton', 'Peel', 'Toronto', 'York')) +
  facet_grid(cols=vars(wave)) +
  xlab('Public Health Unit') +
  ylab('Correlation coefficient') +
  theme_bw()
ggsave('graphs/corr_region_lineplot.png',
       p,
       height = 8.5,
       width = 11,
       units = 'in')
p
```