---
title: "Correlation Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages
```{r}
library(tidyverse)
library(readxl)
library(ggpubr)
library(broom)
```

## Load data
```{r}
# Rt data
o_rt <- read_csv('Estimated_R/ontario_R.csv')
p_rt <- read_csv('Estimated_R/peel_R.csv')
t_rt <- read_csv('Estimated_R/tor_R.csv')
w_rt <- read_csv('Estimated_R/wat_R.csv')
y_rt <- read_csv('Estimated_R/york_R.csv')

# Mobility data
o_mob <- read_excel('Mobility&Rt January 30.xlsx', 
                   sheet = 'Ontario Data')
p_mob <- read_excel('Mobility&Rt January 30.xlsx', 
                   sheet = 'Peel Data')
t_mob <- read_excel('Mobility&Rt January 30.xlsx', 
                   sheet = 'Toronto Data')
w_mob <- read_excel('Mobility&Rt January 30.xlsx', 
                   sheet = 'Waterloo Data')
y_mob <- read_excel('Mobility&Rt January 30.xlsx', 
                   sheet = 'York Data')
```

## Lockdown timeline
On March 17, Premier Ford declared a provincial state of emergency, prohibiting public gatherings larger than 50 people, and ordering the closure of all schools, child care services, libraries, indoor recreation facilities, dine-in bars and restaurants, and all cinemas, theatres, and concert venues.  
Therefore, I will say that the lockdown began in all places on 3/17/20, as they are all cities located in Ontario.  
Ontario: 3/17/20 - 5/14/20  
Peel: 3/17/20 - 6/22/20  
Toronto: 3/17/20 - 6/24/20  
Waterloo: 3/17/20 - 6/12/20  
York: 3/17/20 - 5/19/20  

```{r}
# # take log of R to prevent negative predictions
# mutate(`ln(R)` = log(`Mean(R)`)) %>% 
```

## Function to convert data into useful form
Input: mobility data, rt data, lockdown start date, lockdown end date, columns to keep  
Output: Df of specific region to run regressions on
```{r}
get_data <- function(mob_data, rt_data, lock_start, lock_end, cols) {
  df <- mob_data %>% 
    left_join(rt_data, by=c("date" = "date_start")) %>% 
    select(all_of(cols)) %>% 
    mutate(date = as.Date(date)) %>% 
    rename(retail=retail_and_recreation_percent_change_from_baseline,grocery=grocery_and_pharmacy_percent_change_from_baseline, parks=parks_percent_change_from_baseline, transit=transit_stations_percent_change_from_baseline, workplaces=workplaces_percent_change_from_baseline, residential=residential_percent_change_from_baseline, global=`Global Mobility Score`) %>% 
    mutate(retail_lag7 = lag(retail, 7)) %>% 
    mutate(grocery_lag7 = lag(grocery, 7) ) %>% 
    mutate(parks_lag7 = lag(parks, 7)) %>%
    mutate(transit_lag7 = lag(transit, 7)) %>%
    mutate(workplaces_lag7 = lag(workplaces, 7)) %>%
    mutate(residential_lag7 = lag(residential, 7)) %>%
    mutate(global_lag7 = lag(global, 7)) %>%
    mutate(retail_lag14 = lag(retail, 14)) %>% 
    mutate(grocery_lag14 = lag(grocery, 14) ) %>% 
    mutate(parks_lag14 = lag(parks, 14)) %>%
    mutate(transit_lag14 = lag(transit, 14)) %>%
    mutate(workplaces_lag14 = lag(workplaces, 14)) %>%
    mutate(residential_lag14 = lag(residential, 14)) %>%
    mutate(global_lag14 = lag(global, 14)) %>%
    filter(date>=lock_start & date<=lock_end)
  return(df)
}
```

## Function to get regression data
Input: Df of specific region, desired mobility lag time (0, 7, or 14 days)  
Output: Df of specific region with list of regressions for each mobility
```{r}
get_regressions <- function(data, lag) {
  if(lag==0) {
    retail <- lm(`Mean(R)` ~ retail, data = data)
    grocery <- lm(`Mean(R)` ~ grocery, data = data)
    parks <- lm(`Mean(R)` ~ parks, data = data)
    transit <- lm(`Mean(R)` ~ transit, data = data)
    workplaces <- lm(`Mean(R)` ~ workplaces, data = data)
    residential <- lm(`Mean(R)` ~ residential, data = data)
    global <- lm(`Mean(R)` ~ global, data = data)
  } else if(lag==7) {
    retail <- lm(`Mean(R)` ~ retail_lag7, data = data)
    grocery <- lm(`Mean(R)` ~ grocery_lag7, data = data)
    parks <- lm(`Mean(R)` ~ parks_lag7, data = data)
    transit <- lm(`Mean(R)` ~ transit_lag7, data = data)
    workplaces <- lm(`Mean(R)` ~ workplaces_lag7, data = data)
    residential <- lm(`Mean(R)` ~ residential_lag7, data = data)
    global <- lm(`Mean(R)` ~ global_lag7, data = data)
  } else if(lag==14) {
    retail <- lm(`Mean(R)` ~ retail_lag14, data = data)
    grocery <- lm(`Mean(R)` ~ grocery_lag14, data = data)
    parks <- lm(`Mean(R)` ~ parks_lag14, data = data)
    transit <- lm(`Mean(R)` ~ transit_lag14, data = data)
    workplaces <- lm(`Mean(R)` ~ workplaces_lag14, data = data)
    residential <- lm(`Mean(R)` ~ residential_lag14, data = data)
    global <- lm(`Mean(R)` ~ global_lag14, data = data)
  } else {
    stop('Must enter 0, 7, or 14')
  }

  return(list(retail=retail, grocery=grocery, parks=parks, transit=transit, workplaces=workplaces, residential=residential, global=global))
}
```

## Function to get coefficient data
Input: List of region df's each containing regression info  
Output: Coefficient data df of specified lag time
```{r}
coef_data <- function(data) {
  place <- numeric(length=35)
  mobility <- numeric(length=35)
  coefficient <- numeric(length=35)
  r2 <- numeric(length=35)
  p_val <- numeric(length=35)
  i=1
  j=1
  k=1
  for(loc in data){
    for(reg in loc) {
      place[i] <- names(data)[j]
      mobility[i] <- names(loc)[k]
      coefficient[i] <- tidy(reg)[[2,2]]
      r2[i] <- glance(reg)[[1,2]]
      p_val[i] <- glance(reg)[[1,5]]
      i=i+1
      k=k+1
    }
    j=j+1
    k=1
  }
  df <- data.frame(place, mobility, coefficient, r2, p_val)
  return(df)
}
```

## Pipeline
Columns to keep
```{r}
COLS <- c('sub_region_1', 'date', 'retail_and_recreation_percent_change_from_baseline','grocery_and_pharmacy_percent_change_from_baseline', 'parks_percent_change_from_baseline', 'transit_stations_percent_change_from_baseline', 'workplaces_percent_change_from_baseline', 'residential_percent_change_from_baseline', 'Global Mobility Score', 'Mean(R)')
```

Get data
```{r}
o <- get_data(o_mob, o_rt, '2020-03-17', '2020-05-20', COLS)
p <- get_data(p_mob, p_rt, '2020-03-17', '2020-06-22', COLS)
t <- get_data(t_mob, t_rt, '2020-03-17', '2020-06-24', COLS)
w <- get_data(w_mob, w_rt, '2020-03-17', '2020-06-12', COLS)
y <- get_data(y_mob, y_rt, '2020-03-17', '2020-05-19', COLS)
```

0 day lag
```{r}
zero <- list(
  ontario=get_regressions(o, 0),
  peel=get_regressions(p, 0),
  toronto=get_regressions(t, 0),
  waterloo=get_regressions(w, 0),
  york=get_regressions(y, 0)
)
```

7 day lag
```{r}
seven <- list(
  ontario=get_regressions(o, 7),
  peel=get_regressions(p, 7),
  toronto=get_regressions(t, 7),
  waterloo=get_regressions(w, 7),
  york=get_regressions(y, 7)
)
```

14 day lag
```{r}
fourteen <- list(
  ontario = get_regressions(o, 14),
  peel = get_regressions(p, 14),
  toronto = get_regressions(t, 14),
  waterloo = get_regressions(w, 14),
  york = get_regressions(y, 14)
)
```

Get coefficient data for each time lag  
Add period column, later factor it appropriately  
Combine into one df (total)  
Check if each p-value is statistically significant (<0.01)
```{r}
zero_coef <- coef_data(zero)
zero_coef$period <- rep('Lockdown', times=35)

seven_coef <- coef_data(seven)
seven_coef$period <- rep('7 Days after lockdown', times=35)

fourteen_coef <- coef_data(fourteen)
fourteen_coef$period <- rep('14 Days after lockdown', times=35)

total <- rbind(zero_coef, seven_coef, fourteen_coef)
total$period <- factor(total$period, levels=c('Lockdown', '7 Days after lockdown', '14 Days after lockdown'))
total <- total %>% 
  mutate(
    significant = case_when(
      p_val < 0.01 ~ 'Yes',
      p_val >= 0.01 ~ 'No'
    )
  )

write_csv(total, 'Correlation_Analysis/total.csv')
```

## Barplot
```{r}
p <- ggplot(total, aes(x=place, y=coefficient, fill=mobility)) +
  geom_col(position = 'dodge') +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5)) +
  facet_grid(rows=vars(period)) +
  theme_bw()
ggsave('Correlation_analysis/barplot.png',
       p,
       height = 8.5,
       width = 11,
       units = 'in')
p
```

## Lineplot
```{r}
d <- filter(total, mobility=='global')
p <- ggplot(d, aes(x=period, y=coefficient, group=place, color=place)) +
  geom_point(d, mapping=aes(shape=significant), size=3) +
  geom_line(linetype='dashed') + 
  theme_bw()
ggsave('Correlation_analysis/lineplot.png',
       p,
       height = 8.5,
       width = 11,
       units = 'in')
p
```


